Optimal design of experiment: supercritical fluid extraction case

### 2024-08-06 SFE_poster_DOE/CHISA_2024_Sliczniuk.tex
06.08

\begin{minipage}[b]{0.75\linewidth}
\veryHuge \color{NavyBlue} \textbf{Optimal design of experiment:} \color{Black}\\ % Title
\Huge\textit{supercritical fluid extraction case}\\[2cm] % Subtitle
\huge \textbf{Oliwer Sliczniuk, Pekka Oinas}\\[0.5cm] % Author(s)
\huge School of Chemical Engineering, Aalto University, Finland\\[0.4cm] % University/organization
%\Large \texttt{john@LaTeXTemplates.com} --- 1 (000) 111 1111\\
\end{minipage}
%
\begin{minipage}[b]{0.25\linewidth}
\includegraphics[width=15cm]{Aalto_University_logo.png}\\ \\
\end{minipage}

%\vspace{0.5cm} % A bit of extra whitespace between the header and poster content
\noindent\rule{\textwidth}{3pt}
%\vspace{0.25cm} % A bit of extra whitespace between the header and poster content

@@

\begin{minipage}[b]{0.5\linewidth}
	\begin{itemize}
		\item One-dimensional
		\item Plug flow
		\item No pressure drop
		\item Uniform particle distribution
	\end{itemize}
\end{minipage}
\hfill
\begin{minipage}[b]{0.5\linewidth}
	\begin{itemize}
		\item Two-film theory for a single component
		\item Peng-Robinson equation of state
		\item Decaying extraction kinetic
		\item Empirical correlations
	\end{itemize}
\end{minipage}

@@

\begin{minipage}[h!]{0.33\linewidth}
	\begin{align*}
		c_f &- \text{Solutes concentration in the fluid phase}\\
		c_s &- \text{Solutes concentration in the solid phase}\\
		h~  &- \text{Enthalpy}\\
		y~ &- \text{Extraction yield}\\
		\rho_f &- \text{Density of fluid}\\
		A_f &- \text{Cross-section of the bed}\\
		u~~ &- \text{Darcy velocity}\\
		\phi~~ &- \text{Void fraction}\\
		D_e^M &- \text{Axial mass diffusivity}\\
		D_i^R &- \text{Internal mass diffusivity}
	\end{align*}
\end{minipage}
%
\begin{minipage}[h!]{0.33\linewidth}
	\begin{align*}
		\mu~ &- \text{Particle shape coefficient}\\
		l~ &- \text{Particle length}\\
		\Upsilon~ &- \text{Decaying factor}\\
		\rho_s &- \text{Bulk density of solid bed}\\
		k_m &- \text{Partition factor}\\
		P~ &- \text{Pressure}\\
		T~ &- \text{Temperature}\\
		F~ &- \text{Mass flow rate} \\
		\Sigma~ &- \text{Covariance matrix} \\
		\Theta~ &- \text{Vector of parameters}
	\end{align*}
\end{minipage}

\begin{center}\vspace{0.5cm}
	\includegraphics[width=0.49\linewidth]{Di_Re_F_1.png}
	\includegraphics[width=0.49\linewidth]{Gamma_Re_F_1.png}
	\captionof{figure}{Empirical correlations}
\end{center}%\vspace{1cm}

@@

\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
%------------------------------------------------
%\vspace{1cm}
%\section*{Model-based optimal design of experiment}
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Model-based optimal design of experiment\\}
\end{tcolorbox}

\underline{Fisher information $\mathcal{F}$ (Hessian of the likelihood function) measurs the amount of information} observable random variables carry about a parameters of a distribution that models these variables[3]:

\begin{equation*}
	%\Sigma^{-1} \geq
	\mathcal{F}(t,\Theta; \Xi) = \frac{\partial y(t, \Theta; \Xi)}{\partial \Theta} \Sigma \frac{\partial y(t, \Theta; \Xi)}{\partial \Theta^\top}
\end{equation*}

%The optimal design of experiments involves planning an experiment in such a way that parameters can be estimated without bias and with minimal variance. 
The D-optimality criterion is chosen as the objective function, aiming to \underline{minimize the volume} \underline{of the ellipsoidal confidence region of parameter estimates} under the experimental conditions $\Xi$.

\begin{equation*}
	\begin{aligned} 
		&\Xi^* &= \arg &\min_{ T^{in},~F\in~\Xi}\quad\int_{t_0}^{t_f} - \ln \det \mathcal{F}(t,\Theta; \Xi)~dt  \\
		&\text{subject to}
		& \dot{x} = ~&G~~(x,t,\Theta;\Xi) \\
		&& T^{0} = ~&T^{in}(t=0) \\
		&& 30^\circ C \leq ~&T^{in}(t) ~ \leq 40^\circ C \\
		&& 3.33 \cdot 10^{-5}~\text{kg/s} \leq ~&F~~(t)~ \leq 6.67 \cdot 10^{-5}~\text{kg/s}\\
		&& 100~\text{bar} \leq ~&P~~(t)~ \leq 200~\text{bar} \\
	\end{aligned}
\end{equation*}

This work aims \underline{to improve the precision of the correlation for $D_i^R$} by designing an experiment with dynamically changing operating conditions (\underline{$F$ and $T^{in}$)}. \\

@@

@@

	The system operates for 300 minutes, with a sampling interval of 10 minutes and decision variables adjusted every 15 minutes. Each of the five analysed cases assumes a constant pressure, set at 100, 125, 150, 175, and 200 bar. To identify the global solution, the optimization problem is solved multiple times, each starting from a random initial solution sampled from a uniform distribution. %Figure 2 compares the initial and final values of the cost function across multiple optimization runs for different cases of pressure value.

@@

	The optimal profiles of the inlet temperature and flow rates for each case are showed in Figure 3

	\begin{center}\vspace{0.5cm}
		\includegraphics[width=\linewidth]{Profiles.png}
		\captionof{figure}{Optimal profiles}
	\end{center}%\vspace{1cm}

@@

	\begin{itemize}
		\item The optimal control profiles are similar across all cases.
		\item Low objective values are achieved at pressures near the supercritical point, where variations in the inlet temperature cause significant deviations in the physical properties of CO$_2$ and consequently in the Reynolds number
		\item The mass flow rate is the primary control variable, indicating that the system is more sensitive to mass flow rate changes than to inlet temperature variations.
	\end{itemize}

>>>>

\begin{minipage}[b]{0.75\linewidth}
\veryHuge \color{NavyBlue} \textbf{Optimal design of experiment:} \color{Black}\\ % Title
\Huge\textit{supercritical fluid extraction case}\\[1cm] % Subtitle
\huge \textbf{Oliwer Sliczniuk, Pekka Oinas}\\%[0.5cm] % Author(s)
\huge School of Chemical Engineering, Aalto University, Finland\\[0.4cm] % University/organization
%\Large \texttt{john@LaTeXTemplates.com} --- 1 (000) 111 1111\\
\end{minipage}
%
\begin{minipage}[b]{0.25\linewidth}
\includegraphics[width=12cm]{Aalto_University_logo.png}\\ \\
\end{minipage}

\vspace{-2cm} % A bit of extra whitespace between the header and poster content
\noindent\rule{\textwidth}{3pt}
%\vspace{0.25cm} % A bit of extra whitespace between the header and poster content

@@

\begin{minipage}[b]{0.1\linewidth}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\linewidth}
	\begin{itemize}
		\item One-dimensional
		\item Plug flow
		\item No pressure drop
		\item Uniform particle distribution
	\end{itemize}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\linewidth}
	\begin{itemize}
		\item Two-film theory for a single component
		\item Peng-Robinson equation of state
		\item Decaying extraction kinetic
		\item Empirical correlations
	\end{itemize}
\end{minipage}

@@

\begin{center}\vspace{0.5cm}
	\includegraphics[width=0.49\linewidth]{Di_Re_F.png}
	\hfill
	\includegraphics[width=0.49\linewidth]{Gamma_Re_F.png}
	\captionof{figure}{Empirical correlations}
\end{center}%\vspace{1cm}

@@

%------------------------------------------------
%\vspace{1cm}
%\section*{Model-based optimal design of experiment}
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Model-based optimal design of experiment\\}

%\underline{Fisher information $\mathcal{F}$ (Hessian of the likelihood function) measurs the amount of information} observable random variables carry about a parameters of a distribution that models these variables[3]:

\begin{equation*}
	%\Sigma^{-1} \geq
	\underbrace{\mathcal{F}(t,\Theta; \Xi)}_{\text{Fisher Information}} = \frac{\partial y(t, \Theta; \Xi)}{\partial \Theta} \Sigma \frac{\partial y(t, \Theta; \Xi)}{\partial \Theta^\top}
\end{equation*}

%The optimal design of experiments involves planning an experiment in such a way that parameters can be estimated without bias and with minimal variance. 
The D-optimality criterion is chosen as the objective function, aiming to \underline{minimize the volume} \underline{of the ellipsoidal confidence region of parameter estimates} under the experimental conditions $\Xi$ [3].

\begin{equation*}
	\begin{aligned} 
		&\Xi^* &= \arg &\min_{ T^{in},~F\in~\Xi}\quad\int_{t_0 = 0 \text{min}}^{t_f=300 \text{min}} - \ln \det \mathcal{F}(t,\Theta; \Xi)~dt  \\
		&\text{subject to}
		& \dot{x} = ~&G~~(x,t,\Theta;\Xi) \\
		&& T^{0} = ~&T^{in}(t=0) \\
		&& 30^\circ C \leq ~&T^{in}(t) ~ \leq 40^\circ C \\
		&& 3.33 \cdot 10^{-5}~\text{kg/s} \leq ~&F~~(t)~ \leq 6.67 \cdot 10^{-5}~\text{kg/s}\\
		&& 100~\text{bar} \leq ~&P~~(t)~ \leq 200~\text{bar} \\
	\end{aligned}
\end{equation*}

The aim is \underline{to improve the precision of the correlation for $D_i^R$} by designing an experiment with dynamically changing operating conditions for five cases of constant pressure: 100, 125, 150, 175, and 200 bar. The yield is measured every 10 min and decision variables are adjusted every 15 min.
\end{tcolorbox}

\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Numerics\\}

@@

\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Legend\\}
	\begin{minipage}[b]{0.05\linewidth}
	\end{minipage}
	\hfill
	\begin{minipage}[h!]{0.475\linewidth}
		\begin{align*}
			c_f &- \text{Solutes concentration in the fluid phase}\\
			c_s &- \text{Solutes concentration in the solid phase}\\
			h~  &- \text{Enthalpy}\\
			y~ &- \text{Extraction yield}\\
			\rho_f &- \text{Density of fluid}\\
			A_f &- \text{Cross-section of the bed}\\
			u~~ &- \text{Darcy velocity}\\
			\phi~~ &- \text{Void fraction}\\
			D_e^M &- \text{Axial mass diffusivity}\\
			D_i^R &- \text{Internal mass diffusivity}
		\end{align*}
	\end{minipage}
	%
	\begin{minipage}[h!]{0.475\linewidth}
		\begin{align*}
			\mu~ &- \text{Particle shape coefficient}\\
			l~ &- \text{Particle length}\\
			\Upsilon~ &- \text{Decaying factor}\\
			\rho_s &- \text{Bulk density of solid bed}\\
			k_m &- \text{Partition factor}\\
			P~ &- \text{Pressure}\\
			T~ &- \text{Temperature}\\
			T^{in} &- \text{Inlet temperature}\\
			F~ &- \text{Mass flow rate} \\
			\Sigma~ &- \text{Covariance matrix} \\
			\Theta~ &- \text{Vector of parameters}
		\end{align*}
	\end{minipage}
\end{tcolorbox}

@@

	To identify the global solution, the optimization problem is solved multiple times, each starting from a random initial solution sampled from a uniform distribution. %Figure 2 compares the initial and final values of the cost function across multiple optimization runs for different cases of pressure value.

@@

	The optimal profiles of the controls and the yield curves are showed below

	\begin{center}\vspace{0.5cm}
		\includegraphics[width=0.92\linewidth]{Profiles_all.png}
		\captionof{figure}{Optimal profiles}
	\end{center}%\vspace{1cm}

@@

	\begin{itemize}
		\item The lowest values of the cost function are achieved near the supercritical point, where variations in the inlet temperature cause significant deviations in the physical properties of CO$_2$ and $Re$
		\item The optimal profiles of the controls are similar across all cases
		\item The mass flow rate is the primary control variable, indicating that the system is more sensitive to mass flow rate changes than to inlet temperature variations
		\item The optimal yield profiles are characterized by waving behaviour 
		\item Experiments which result in the highest yields are not necessary the most informative
	\end{itemize}
